### 2024-12-01 paper/paper.qmd
Update paper

---
title: "Factors influencing Total Income in the United States"
subtitle: "Analyzing Factors influencing Total Income in the United States of America from 2019 to 2022"
author: 
  - Wen Han Zhao
thanks: "Code and data are available at: [https://github.com/younazhao/Income-Influencing-Factors.git](https://github.com/younazhao/Income-Influencing-Factors.git)."
date: 1 December 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

```{r}
#| include: false
#| warning: false
#| message: false
#install.packages("kableExtra")
library(tidyverse)
library(arrow)
library(dplyr)
library(tibble)
library(knitr)
library(kableExtra)
```

```{r}
#| include: false
#| warning: false
#| message: false

data <- read_parquet(here::here("data/02-analysis_data/analysis_data.parquet"))
```

# Introduction

Income inequality is a central concern in economic research, reflecting disparities in access to opportunities, resources, and outcomes. In the United States, total income is influenced by a variety of factors, including personal, professional, and demographic characteristics. This paper explores the relationship between total income and predictors such as marital status, sex, educational level, labor force participation, and occupation. By employing statistical modeling techniques, we aim to uncover the extent to which these variables shape income levels and contribute to broader income disparities.

The estimand of this study is the expected total income for individuals based on their social, personal and occupational characteristics. Specifically, the analysis seeks to quantify how factors such as marital status, sex, educational attainment, participation in the labor force, and occupational category predict variations in total income across the U.S. population. This includes identifying both independent and among these variables.

Preliminary results indicate that educational attainment is one of the strongest predictors of total income, with higher levels of education yielding significantly greater earnings. Marital status also plays a notable role, with married individuals generally earning more than their single counterparts. Gender disparities in income persist, with men earning more on average than women across most occupational categories. Labor force participation and occupation further highlight critical differences, as full-time workers and those in higher-skilled professions consistently report higher incomes.

Understanding the drivers of income disparities is essential for designing policies that promote economic equity and mobility. By identifying how marital status, gender, education, labor force participation, and occupation contribute to income levels, this research offers actionable insights for policymakers, employers, and educators. Addressing these disparities not only fosters greater fairness but also enhances the overall economic productivity and social cohesion of the United States.

To do this, total income was analyzed in Section 2 @sec-data. The prediction from the model has show accuracy in lower total income group. The model has shown difficulties in predicting higher total income. Section 3 @sec-model has further explanation on the Bayesian Generalized Linear Model. Section 4 @sec-results show the coefficient beta from the model. Section 5 @sec-discussion focus on the weakness of this study and further possible improvements. Lastly, Section 6 @sec-appendix contains methodology to perform a survey to collect data for Total Income.

# Data {#sec-data}

## Overview {#sec-data-overview}

The Survey of Consumer Finances (SCF), conducted in 2022 covering the period 1989 to 2022 by the Board of Governors of the Federal Reserve System of the United States [@scf2022], provides a comprehensive snapshot of household financial conditions across the nation. This dataset includes detailed information related to total income, capturing key predictors such as sex, education level, occupation category, salary income, and income from other sources. By offering rich, granular data, the SCF serves as an invaluable resource for analyzing the factors influencing income distribution and economic disparities within the United States.

## Measurement {#sec-data-measurement}

The survey contained a panel element over 2 periods. Respondents to the 1983 survey were re-interviewed in 1986 and 1989. Respondents to the 2007 survey were re-interviewed in 2009. In order to ensure the representativeness of the study, respondents are selected randomly in order to attempt families from all economic strata.

## Outcome variables{#sec-data-outcome}

The histogram represents the distribution of our outcome variables total income among survey participants. From the graph, we can see that the graph shows slightly right skewed but mostly symmetric by following a bell shape curve. It shows a mean centered 11 to 12, indicating that most individuals in the survey have incomes within this range. The spread of the graph shows approximately from 8 to 16 of the log income and contains only a few extreme outliers at the tails. Overall, the distribution of log income, the outcome variable, follows a normal distribution.

```{r}
#| label: fig-hist
#| fig-cap: Distribution of Total Income 
#| echo: false
data$log_income <- log(data$income)
ggplot(data, aes(x = log_income)) +
  geom_histogram(binwidth = 0.5, fill = "blue", color = "black", alpha = 0.7) +
  labs(x = "Log Income", y = "Frequency") +
  theme_minimal()
```

Quantile-quantile plot is included to compare the log-transformed income data' distribution to a theoretical normal distribution. The data should expect to follow closely to the blue dashed line to indicate a normal distribution. From the graph, we can see that the middle of the plot lies closely to the line which indicate that the data follows a normal distribution. The right tail have various points deviate above the line, which suggest a heavier right tail. The deviation form the right tail is expected, since higher income are more random and harder to predict.

```{r}
#| label: fig-qq
#| fig-cap: Comparison of Income Distribution using QQ Plot
#| echo: false
#| warning: false
#| message: false

ggplot(data = data, aes(sample = log_income)) +
  geom_qq(color = "black", size = 1, alpha = 0.7) + # Adds blue points with some transparency
  geom_qq_line(color = "blue", linetype = "dashed", size = 1) + # Adds a red dashed reference line
  theme_minimal() +
  labs(
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  )

```

## Predictor variables{#sec-data-predictor}

Predictors variables selected are Age, Sex, Marital Status, Education Level, Labour Force, Work Status, and Occupation category.

```{r}
#| label: fig-pred-age
#| fig-cap: Log Income by Age
#| echo: false
#| warning: false
#| message: false

plot(data$Age, data$log_income, type = "p", 
     xlab = "Age", ylab = "Log Income")
abline(lm(log_income ~ Age, data = data), col = "red", lwd = 2)

```

```{r}
#| label: fig-pred-sex
#| fig-cap: Log Income by Sex
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = Sex, y = log_income)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  xlab("Labour Force") +
  ylab("Log Income") +
  theme_minimal()

```

```{r}
#| label: fig-pred-married
#| fig-cap: Log Income by Marital Status
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = Marital_status, y = log_income)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  xlab("Labour Force") +
  ylab("Log Income") +
  theme_minimal()

```

```{r}
#| label: fig-pred-educ
#| fig-cap: Log Income by Educational Level
#| echo: false
#| warning: false
#| message: false

# Simplify the education levels
data$Education_simplified <- case_when(
  data$educ == -1 ~ "Less than 1st grade",
  data$educ == 1 ~ "1st-4th grade",
  data$educ == 2 ~ "5th-6th grade",
  data$educ == 3 ~ "7th-8th grade",
  data$educ == 4 ~ "9th grade",
  data$educ == 5 ~ "10th grade",
  data$educ == 6 ~ "11th grade",
  data$educ == 7 ~ "12th grade, no diploma",
  data$educ == 8 ~ "High school graduate",
  data$educ == 9 ~ "Some college",
  data$educ == 10 ~ "Associate degree (vocational)",
  data$educ == 11 ~ "Associate degree (academic)",
  data$educ == 12 ~ "Bachelor's degree",
  data$educ == 13 ~ "Master's degree",
  data$educ == 14 ~ "Professional/Doctorate"
)
# Reorder the 'Education_simplified' based on 'educ' numbers from small to large
data$Education_simplified <- factor(data$Education_simplified, 
                                    levels = c("Less than 1st grade", "1st-4th grade", "5th-6th grade", 
                                               "7th-8th grade", "9th grade", "10th grade", "11th grade", 
                                               "12th grade, no diploma", "High school graduate", "Some college", 
                                               "Associate degree (vocational)", "Associate degree (academic)", 
                                               "Bachelor's degree", "Master's degree", 
                                               "Professional/Doctorate"))

# Create the plot with simplified labels in correct order
ggplot(data, aes(x = Education_simplified, y = log_income)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  xlab("Education Level") +
  ylab("Log Income") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for better readability

```

```{r}
#| label: fig-pred-lf
#| fig-cap: Log Income by Labour Force
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = Labor_force, y = log_income)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  xlab("Labour Force") +
  ylab("Log Income") +
  theme_minimal()

```

```{r}
#| label: fig-pred-worksta
#| fig-cap: Log Income by Work Status
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = Work_status_category, y = log_income)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  xlab("Work Status") +
  ylab("Log Income") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
#| label: fig-pred-occup
#| fig-cap: Log Income by Occupation Category
#| echo: false
#| warning: false
#| message: false

ggplot(data, aes(x = Occupation_Classification, y = log_income)) +
  geom_boxplot(fill = "lightblue", color = "darkblue") +
  xlab("Occupation Category") +
  ylab("Log Income") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
\pagebreak

# Model{#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up{#sec-model-setup}

Define $y_i$ as the Total Income for each individual participated in the survey in the United States at 2022. The predictors are categorical variables with different levels explained in @sec-data-.

\begin{align*}
\log(y_i) &= \beta_0 
    + \beta_1 \cdot \text{Sex} 
    + \beta_2 \cdot \text{Education Level} 
    + \beta_3 \cdot \text{Marital Status} \\
&\quad + \beta_4 \cdot \text{Labor Force} 
    + \beta_5 \cdot \text{Work Status Category} 
    + \beta_6 \cdot \text{Occupation Classification} 
    + \epsilon, \\
\beta_0              &\sim \text{Normal}(10, 5), \\
\beta_1, \beta_2, \ldots, \beta_6 &\sim \text{Normal}(0, 2.5), \\
\epsilon             &\sim \text{Normal}(0, \delta^2), \\
\delta               &\sim \text{Exponential}(1).
\end{align*}

## Linear Model {#sec-model-lm}

We have started with the linear model for predicting Total Income based on the selected variables. The model set up is:

\begin{align*}
\hat{\text{income}} &= \beta_0 + \beta_1 \cdot \text{Sexmale} + \beta_2 \cdot \text{Education Level} + \beta_3 \cdot \text{Marital Status} + \beta_4 \cdot \text{Labor Force} \\
&\quad + \beta_5 \cdot \text{num\_work} + \beta_6 \cdot \text{num\_occu}
\end{align*}

```{r}
#| label: fig-lm
#| fig-cap: Linear Model has limited predictive accuracy for Total Income
#| echo: false
#| warning: false
#| message: false
lm_model <- lm(formula = income ~ Sex + educ + married + lf + num_work + num_occu, data = data)
prediction_1<- predict(lm_model, data)

ggplot(data, aes(x = income, y = prediction_1)) +
  geom_point(color = 'blue') +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "red") +
  labs(title = "Actual vs Predicted Values",
       x = "Actual pct",
       y = "Predicted pct") +
  theme_minimal()

```

```{r}
#| label: tbl-lmresult
#| fig-cap: Table for Linear Model Results
#| echo: false
#| warning: false
#| message: false

# Table for Linear Model Results
table_description <- tibble(
  Variable = c("(Intercept)", "Sexmale", "educ", "married", "lf", "num_work", "num_occu", 
               "Residual Standard Error", "Multiple R-squared", "Adjusted R-squared", "F-statistic", "p-value"),
  Estimate = c(-348940, 668026, 227466, -1103395, 116525, 1892834, -1273936, 
               "1236000", "0.02223", "0.02198", "86.3", "< 2e-16"),
  Std_Error = c(870243, 263287, 33274, 234718, 353617, 144769, 123257, 
                "", "", "", "", ""),
  t_value = c(-0.401, 2.537, 6.836, -4.701, 0.330, 13.075, -10.336, 
              "", "", "", "", ""),
  Pr_gt_Abs_t = c(0.6884, 0.0112, 8.35e-12, 2.60e-06, 0.7418, "< 2e-16", "< 2e-16", 
                  "", "", "", "", ""),
  Significance = c("", "*", "***", "***", "", "***", "***", "", "", "", "", "")
)

# Draw the table 
kable(table_description, col.names = c("Variable", "Estimate", "Std. Error", "t-value", "Pr($\\gt \\mid t \\mid $)", "Significance")) |>
  kable_styling(latex_options = c("striped", "hold_position"))
```

## Model justification{#sec-model-justification}
We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`. The prior intercept have been set to $N(10,5)$ because the histogram @fig-hist indicates a mean near 10-12 with variation of 5. In addition, various predictors would have strong correlation with our outcome variable Total Income. By setting the prior intercept would capture well the additional knowledge compared to the default priors from `rstanarm` package.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-readRDS(file = here::here("models/bayesian_model_2.rds"))
#model plot
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary::modelsummary(list("First model" = first_model),statistic = "mad",fmt = 2)
```

# Discussion{#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
#  theme_classic() +
#  theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

\newpage

# References